---
title: "Factor Model"
author: "By 21031"
date: "2021/12/21"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Factor Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

__StatComp21031__ is a simple R package developed to compare the performance of R and R++ (implemented through the R package _Rcpp_) for the 'Statistical Computing' course. Two functions are considered, namely, _FM_ (Determine the number of factors in approximate factor models) and _DFM_ (Determine the Number of Factors in the general dynamic factor model). 

The R package 'microbenchmark' can be used to benchmark the above _FM_ and _DFM_ functions.

## Benchmarking _FM_ and _DFM_

The \code{FM} function is adapted by Bai & Ng method to minimize information criterion with two penalties in static factor models.

$$
\widehat{K}=\arg \min _{0 \leq K_{1} \leq M} \log \left\{\frac{1}{p n}\left\|\mathbf{Y}-n^{-1} \mathbf{Y} \widehat{\mathbf{F}}_{K_{1}} \widehat{\mathbf{F}}_{K_{1}}^{\prime}\right\|_{F}^{2}\right\}+K_{1} g(n, p)
$$
where $M$ is a prescribed upper bound, $\widehat{\mathbf{F}}_{K_{1}}$ is a $n \times K_{1}$ matrix whose columns are $\sqrt{n}$ times the eigenvectors corresponding to the $K_{1}$ largest eigenvalues of the $n \times n$ matrix $\mathbf{Y}^{\prime} \mathbf{Y}$;

Here we choose two penalties below.

$$
\mathrm{IC} 1: g(n, p)=\frac{p+n}{p n} \log \left(\frac{p n}{p+n}\right)
$$

$$
\text { IC2 }: g(n, p)=\frac{p+n}{p n} \log \min \{p, n\}
$$

The \code{DFM} function is used to minimize information criterion with two penalties in the general dynamic factor model. Comparing to the static factor model, parameters of information criterion in this model is not fixed, changing the fixed p,n and c to vectors.

The source R code for _FM_ is as follows:
```{r,eval=FALSE}
FM<-function (Y) 
{
  p = nrow(Y)
  n = ncol(Y)
  Y <- Y - t(t(apply(t(Y), 2, mean))) %*% matrix(1, 1, n)
  rmax = 10
  IC = array(0, c(2, rmax))
  frob = array(0, c(rmax))
  for (k in 1:rmax) {
    V <- eigen(t(Y) %*% Y)$vectors
    V = as.matrix(V)
    Dd <- eigen(t(Y) %*% Y)$values
    Dd = as.vector(Dd)
    #eigenvectors corresponding to the K largest eigenvalues of Y'Y
    F <- V[, 1:k]
    LamPCA = Y %*% F/n
    uhat = Y - LamPCA %*% t(F)
    #two penalty
    frob[k] = sum(diag(uhat %*% t(uhat)))/(p * n)
    first = log((p * n)/(p + n)) * (p + n)/(p * n)
    second = log(min(p, n)) * (p + n)/(p * n)
    IC[1, k] = log(frob[k]) + k * first
    IC[2, k] = log(frob[k]) + k * second
  }
  K1 = which.min(IC[1, ])
  K2 = which.min(IC[2, ])
  result <- list(K1 = K1, K2 = K2, IC = IC)
  return(result)
}
```

The source R code for _DFM_ is as follows:

```{r,eval=FALSE}
DFM<-function (Y) 
{
  p = nrow(Y)
  n = ncol(Y)
  Y <- Y - t(t(apply(t(Y), 2, mean))) %*% matrix(1, 1, n)
  c = seq(0.05, 5, length = 100)
  re = 20
  rmax = 10
  IC = array(0, c(2, re, rmax, 100))
  first = array(1, c(re))
  second = array(1, c(re))
  pi = array(1, c(re))
  ni = array(1, c(re))
  for (i in 1:re) {
    pi[i] = min(i * floor(p/re) + min(p, 5), p)
    ni[i] = min(i * floor(n/re) + min(n, 5), n)
    if (i == re) {
      pi[i] = p
      ni[i] = n
    }
    Yi = Y[1:pi[i], 1:ni[i]]
    frob = array(0, c(rmax))
    for (k in 1:min(pi[i], ni[i], rmax)) {
      V <- eigen(t(Yi) %*% Yi)$vectors
      V = as.matrix(V)
      Dd <- eigen(t(Yi) %*% Yi)$values
      Dd = as.vector(Dd)
      F <- V[, 1:k]
      LamPCA = Yi %*% F/ni[i]
      uhat = Yi - LamPCA %*% t(F)
      frob[k] = sum(diag(uhat %*% t(uhat)))/(pi[i] * ni[i])
      first[i] = log((pi[i] * ni[i])/(pi[i] + ni[i])) * 
        (pi[i] + ni[i])/(pi[i] * ni[i])
      second[i] = log(min(pi[i], ni[i])) * (pi[i] + ni[i])/(pi[i] * 
                                                             ni[i])
      for (l in 1:100) {
        IC[1, i, k, l] = log(frob[k]) + c[l] * k * first[i]
        IC[2, i, k, l] = log(frob[k]) + c[l] * k * second[i]
      }
    }
  }
  rhat = array(0, c(2, re, 100))
  for (i in 1:re) {
    for (l in 1:100) {
      m = min(pi[i], ni[i], rmax)
      temp1 = which.min(IC[1, i, 1:m, l])
      rhat[1, i, l] = temp1
      temp2 = which.min(IC[2, i, 1:m, l])
      rhat[2, i, l] = temp2
    }
  }
  Sc1 = array(0, c(100))
  Sc2 = array(0, c(100))
  for (l in 1:100) {
    # i using sd, with l fixed
    Sc1[l] = sd(rhat[1, , l])
    Sc2[l] = sd(rhat[2, , l])
  }
  c1vec = which(Sc1 == 0)
  ctemp1 = c1vec[1]
  K1 = rhat[1, 1, ctemp1]
  
  c2vec = which(Sc2 == 0)
  ctemp2 = c2vec[1]
  K2 = rhat[2, 1, ctemp2]
  
  result <- list(K1 = K1, K2 = K2)
  return(result)
}
```

The R code for benchmarking _FM_ and _DFM_ is as follows.

```{r,eval=TRUE}
library(StatComp21031)
library(microbenchmark)
p=100
n=100
Y<-array(rnorm(p*n),dim=c(p,n))
tm2 <- microbenchmark(
  K1 = FM(Y),
  K2 = DFM(Y)
)
knitr::kable(summary(tm2)[,c(1,3,5,6)])
```

The results show _DFM_ computes slower than the _FM_ which is obvious because _DFM_ include higher dimensional parameters than _FM_.

